---
title: "Wrangling specimen occurrence data"
author: "Alice Armstrong"
date: "2023-03-06"
output: html_document
---



This script takes georeferenced data for Nile crcodiles and wrangles these for later analysis.


#Load packages# 
```{r}
library(gridExtra)
library(sf)
library(patchwork)
library(tidyverse)
```


#Load raw data from GBIF and NHM#
```{r}
NHM<-read.csv('croc_neat.csv')
GBIF<-read.csv('GBIF_final.csv')

```


#Select columns of interest#
```{r}
GBIF1<-GBIF %>% select('Date','Long', 'Lat','Extent','Cer' )
NHM1<-NHM %>% select('Date','Long', 'Lat','Extent','Cer', 'Keep' )
```

#Drop NAs#
```{r}
GBIF2<-drop_na(GBIF1)
NHM2<-drop_na(NHM1)
```

#Remove rows listed as ‘N’ in NHM data (these were records with unclear information/ other issues)#
```{r}
NHM3<-subset(NHM2, Keep == "Y")
```

#Rename NHM and GBIF data, and limt NHM column names to match GBIF#
```{r}
fullNHM<-NHM3%>%select ('Date','Long', 'Lat','Extent', 'Cer')
fullGBIF<-GBIF2

```

#Merge datasets#
```{r}
fulldata<-rbind(fullNHM,fullGBIF)
nrow(fulldata)
```


#Make dates appear in descending order#
```{r}
fulldata<-fulldata%>% arrange(desc(Date))
```

#Filter for extents same as or less than 5000km#
```{r}
fulldata<-fulldata%>% filter(Extent<=5000)
```

#Filter for certainty scores the same as or greater than 50#
```{r}
fulldata<-fulldata%>% filter(Cer>=50)
```

#Omit data from the last 10 years (so we only have historical data)#
```{r}
fulldata<-fulldata%>% filter(Date<=1999)
```

#Write file#
```{r}
write.csv(fulldata, "full_specimin_localility",row.names = T)
```