---
title: "Plotting specimen occurrence data"
author: "Alice Armstrong"
date: "2023-03-06"
output: html_document
keep_md: yes

This script allows you to visualize georeferenced Nile crocodile specimens from GBIF and the NHM (London)

---
```{r}
library(gridExtra)
library(sf)
library(dplyr)
library(patchwork)
library(tidyverse)

```


#Load raw data from GBIF and NHM#
```{r}
NHM<-read.csv('croc_neat.csv')
GBIF<-read.csv('GBIF_final.csv')
```

```{r}
GBIF1<-GBIF %>% select('Date','Long', 'Lat','Extent','Cer' )
NHM1<-NHM %>% select('Date','Long', 'Lat','Extent','Cer', 'Keep' )
```

#Drop NAs#
```{r}
GBIF2<-drop_na(GBIF1)
NHM2<-drop_na(NHM1)
```

#Remove rows listed as ‘N’ in NHM data (these were records with unclear information/ other issues)#
```{r}
NHM3<-subset(NHM2, Keep == "Y")
```

#Rename NHM and GBIF data, and limt NHM column names to match GBIF#
```{r}
fullNHM<-NHM3%>%select ('Date','Long', 'Lat','Extent', 'Cer')
fullGBIF<-GBIF2
```

#filter for extents same as or less than 5000km#
```{r}
fullNHM2<-fullNHM%>% filter(Extent<=5000)
fullGBIF2<-fullGBIF%>% filter(Extent<=5000)
```

#Filter for certainty scores the same as or greater than 50#
```{r}
fullNHM2<-fullNHM%>% filter(Cer>=50)
fullGBIF2<-fullGBIF%>% filter(Cer>=50)
```

#Filter years for NHM (1850:1900,1901:1951,1952:2002)#
```{r}
fiftyyrs1 <- fullNHM2 %>% filter(Date %in% (1850:1900))
fiftyyrs2 <- fullNHM2 %>% filter(Date %in% (1901:1951))
fiftyyrs3 <- fullNHM2 %>% filter(Date %in% (1952:2002))
```

#Filter years for GBIF (1850:1900,1901:1951,1952:2002)#
```{r}
yrs1 <- fullGBIF2 %>% filter(Date %in% (1850:1900))
yrs2 <- fullGBIF2 %>% filter(Date %in% (1901:1951))
yrs3 <- fullGBIF2 %>% filter(Date %in% (1952:2002))
```

#Helper functions for plotting#
```{r}
remove_y <-
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
remove_x <-
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

#convert the NHM coordinates into spatial points then sf format#
```{r}

fiftyyrs11 <-
  fiftyyrs1 %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

fiftyyrs22 <-
  fiftyyrs2 %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

fiftyyrs33 <-fiftyyrs3 %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

```

#convert the GBIF coordinates into spatial points then sf format#
```{r}
yrs11 <-
  yrs1 %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

yrs22 <-
  yrs2 %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

yrs33 <-
  yrs3 %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)
```

#Make a base map of the land#
```{r}
baseMap <-
  rnaturalearth::ne_countries(returnclass = 'sf')
```

# Plot base map#
```{r}
ggplot(baseMap) +
  geom_sf(fill = "lightsteelblue1", lwd = 0.4) +
  theme_void()
```

# Choose coordinates to limit to Africa#
```{r}
africa_bbox <- c(-14.414063, -37.996163,
                 53.613281, 27.994401)
xlim_af <- c(africa_bbox[1], africa_bbox[3])
ylim_af <- c(africa_bbox[2], africa_bbox[4])
```

#Plot NHM and GBIF data from 1850-1900#
```{r}
map1<-ggplot(baseMap) +
  geom_sf(fill = "aliceblue", lwd = 0.4) +
  # Add species
  geom_sf(data = fiftyyrs11, colour = "blue", size = 3, alpha=0.5) +
  # Add equator line
  geom_hline(aes(yintercept = 0), colour = "black") +
  geom_sf(data = yrs11, colour = "red", size = 2, alpha=0.5) +
  coord_sf(xlim = xlim_af, ylim = ylim_af, expand = TRUE) +
  theme_minimal() +
  remove_x + # gets rid of x axis as we don't need it
  remove_y # gets rid of y axis

```

# Plot NHM and GBIF data from 1901-1951#
```{r}
map2<-ggplot(baseMap) +
  geom_sf(fill = "aliceblue", lwd = 0.4) +
  # Add species
  geom_sf(data = fiftyyrs22, colour = "blue", size = 3, alpha=0.5) +
  # Add equator line
  geom_hline(aes(yintercept = 0), colour = "black") +
  geom_sf(data = yrs22, colour = "red", size = 2, alpha=0.5) +
  coord_sf(xlim = xlim_af, ylim = ylim_af, expand = TRUE) +
  theme_minimal() +
  remove_x + # gets rid of x axis as we don't need it
  remove_y 
```

#Plot NHM and GBIF data from (1952:2002#
```{r}
map3<-ggplot(baseMap) +
  geom_sf(fill = "aliceblue", lwd = 0.4) +
  # Add species
  geom_sf(data = fiftyyrs33, colour = "blue", size = 3, alpha=0.5) +
  # Add equator line
  geom_hline(aes(yintercept = 0), colour = "black") +
  geom_sf(data = yrs33, colour = "red", size = 2, alpha=0.5) +
  coord_sf(xlim = xlim_af, ylim = ylim_af, expand = TRUE) +
  theme_minimal() +
  remove_x + # gets rid of x axis as we don't need it
  remove_y # gets rid of y axis

```


#plot maps side by side#
```{r}
specimindata<-map1 + map2 + map3
```

#label maps for figure (i,ii)#
```{r dev = "jpg"}
specimindata<-specimindata + plot_annotation (tag_levels= 'i')
```


