---
title: "Maxent"
author: "Alice Armstrong"
date: "2023-03-06"
output: html_document
---

This script runs a Maxent species distribution model on the historical Nile crocodile specimen data from GBIF and the NHM.

#Load packages#
```{r}
library(raster)
library(tidyverse)
library(dismo)
library(terra)
library(geodata)
library(sf)
library(sp)
library(rgdal)
library(sfheaders)
library(rnaturalearth)
library(geodata)
library(reshape2)
library(rJava)
library(maps)
library(mapdata)
library(maptools)
library(jsonlite)
library(usdm)
library(rgeos)
library(spThin)
library(ids)
```

#set seed to correct for randomness#
```{r}
set.seed(1)
```

##Load the Wordclim data#
```{r}
geodata::worldclim_global 
modelEnv<- worldclim_global(var='bio', res=10, path='data')
```

#stack spatrasters#
```{r}
modelEnv<-raster::stack(modelEnv) 
```

# Relabel the variables#
```{r}
bioclim_names <- paste0('bio', 1:19)
names(modelEnv) <- bioclim_names
```
#Check VIF to deal with multicollinearity #
```{r}
usdm::vifstep(modelEnv,th=3)
```

# remove predictors with VIF values greater than 3#
```{r}
modelEnv=dropLayer(modelEnv, 
  c("bio1", "bio4", "bio5", "bio6", "bio7", "bio8", "bio9", "bio10",
    "bio11","bio12", "bio13","bio16","bio17"))
```

#load and edit locality data#
```{r}
data<-read.csv('full_specimin_localility')
data<-data%>%dplyr::select('Long','Lat','Extent','Cer')
raw.data<-data%>%dplyr::select('Long','Lat')
```

#correct for potential spatial auto correlation#


#check grid limits#
```{r}
max(raw.data$Long) #55.46
min(raw.data$Long) # -14.76935
max(raw.data$Lat) #29.25896
min(raw.data$Lat) # -34.00129

```

#set grid limits#
```{r}
x <- seq(from = -20, to =60, by = 0.16666)
y <- seq(from = -40, to = 35, by = 0.16666)
xy <- expand.grid(x, y)
colnames(xy)<-c("Long","Lat")

bounds <- data.frame(xl = (xy$Long - 0.083333333),
                     xu = (xy$Long + 0.083333333), 
                     yl = (xy$Lat - 0.083333333), 
                     yu = (xy$Lat + 0.083333333),
                     xy)
bounds$id<-random_id(216931, 4)
```

#set for loop that takes one occurrence per grid cell#
```{r}
for (i in 1:216931){
  grid <- bounds[i,] 
  for (j in 1:233){ 
    x <- raw.data[j,1] 
    y<- raw.data[j,2]
    if (grid$xl <= x & grid$xu >= x &  grid$yl <= y & grid$yu >= y) {
      raw.data[j,4:6]<-grid[,5:7]
    } 
  } 
}

```

#take each unique record from the resulting dataset#
```{r}
unique.species<-unique(raw.data[,4:6]) 
data<-unique.species
data<-data%>%dplyr::select('Long.1', 'Lat.1')
```


#make model extent#
```{r}
model.extent<-extent(min(data$Long)-10,max(data$Long)+10,
                     min(data$Lat)-10,max(data$Lat)+10)
```

#crop data to local range#
```{r}
modelEnv<-crop(modelEnv,model.extent)

```

#withold 20% of sample for testing#
```{r}
fold <- kfold(data, k=5)
occtest <- data[fold == 1, ]
occtrain <- data[fold != 1, ]

```

#fit maxent#
```{r}
maxent_model<-maxent(modelEnv,occtrain)
```

#plot maxent#
```{r}
plot(maxent_model)
```

#predict on entire dataset#
```{r}
maxent_entire <- predict(maxent_model, modelEnv)

```

#plot#
```{r}
plot(maxent_entire)

```

#response curves#
```{r}
par(mar = c(1, 1, 1, 1))
response(maxent_model)
```

#load background data (make sure raster is stacked)#
```{r}
modelEnv<-raster::stack(modelEnv) 
bg <- randomPoints(modelEnv,10000)
```

#evaluate/ test data#
```{r}
model_evaluation <- dismo::evaluate(maxent_model, p=occtest, a=bg, x=modelEnv)
```

#load ICUN polygons#
```{r}
iucn <- 
  st_read(dsn = "~/R/CROCODILES_ALLIGATORS") %>%
  st_as_sf() %>%
  st_transform(4326)
```

#Select only Nile Crocs#
```{r}
iucn2<-subset(iucn, binomial == "Crocodylus niloticus")
```

#make sure sf objects are the same projection#
```{r}
iucn2<- st_transform(iucn2, 4326)
```

#plot with ICUN polygons#
```{r}
par(mar=c(0,0,0,0))
plot(maxent_entire)
plot(st_geometry(iucn2), add=TRUE, lwd=2)
```

#calculate the number of points in raster with highest habitat suitability (1.0-0.6)#
```{r}
historical <- length(which(values(maxent_entire) 
                           <= 1.0 & values(maxent_entire) >= 0.6))
```

#crop raster to iucn points#
```{r}
inter<-intersect(maxent_entire,iucn2)
plot(inter)
```

#calculate the number of iucn points in the cropped raster with highest habitat suitability#
```{r}
present_day<-sum(values(inter)==1.0,na.rm=TRUE)
present_day <- length(which(values(inter) <= 1.0 & values(inter) >= 0.6))
```

#calculate percentage overlap of suitable habitat#
```{r}
overlap<-present_day/historical
overlap<-overlap*100
```

