---
title: "Population extractions"
author: "Alice Armstrong"
date: "2023-03-09"
output: html_document
---

#Load packages#
```{r setup, include=FALSE}
library(sf)
library(sfe)
library(tidyverse) 
library(raster) 
library(rgdal)
library(here)
```

#Load custom function#
```{r}
source(here::here("04 Functions anthropocene data.R"))
```

#Read in the data and set coordinate ref system. 
#For some reason st_read makes everything a factor 
#so convert Extent etc back to numeric first and relevel redlist
```{r}
specs_errors <- 
  st_read(here("crocs-extents.csv")) %>%  
  st_set_crs(4326)
```

#Define root folder and then folders within this#
```{r}
root <- "Population_high"
folders <- list.files(here(root))
paths <- paste0(root, "/", folders)

```

#Extract list of files in all the folders with paths attached#
```{r}
files<- NULL
for (i in 1:length(paths)){
  files <- c(files, paste0(paths[i], "/", list.files(here(folders[i]))))
}   

```

#Extract list of files in all the folders#
```{r}
filenames <- NULL
for (i in 1:length(paths)){
  filenames <- c(filenames, list.files(here(folders[i])))
} 
```

#Remove .asc from filenames#
```{r}
filenames <- str_remove(filenames, ".asc")
```

#Read in all the data#
```{r}
for (i in 1:length(files)){
  assign(filenames[i], raster(here(files[i]), 
                              crs = "+proj=longlat +datum=WGS84"))}
```

#And extract the HPD...#
```{r}
overlaps_hpd <- extract_hpd(crocs_errors)
                             
```

## Write to file#
```{r}
st_write(overlaps_hpd, here("data/overlaps_hpd.csv"))
```
