---
title: "Areas of overlap between historic and current ranges"
author: "Alice Armstrong"
date: "2023-03-09"
output: html_document
---
#Load packages#
```{r}
devtools::install_github("JCur96/sfe", force = TRUE)
library(sfe)
library(rgeos)
library(rgdal)
library(Hmisc)
library(tidyverse)
library(here)
library(patchwork)
library(knitr)
library(rnaturalearth)
library(ggplot2)
library(devtools)
```


#Croc georeferenced data with extents#
```{r}
nhm <- st_read("crocs-points.csv")
nhm <- nhm %>%
  st_set_crs(4326)

```

#Read in the IUCN range maps#
```{r}
iucn <- 
  st_read(dsn = "~/R/CROCODILES_ALLIGATORS") %>%
  st_as_sf() %>%
  st_transform(4326)

```

#Select only Nile crocs#
```{r}
iucn2<-subset(iucn, binomial == "Crocodylus niloticus")
```

#load basemap#
```{r}
baseMap <-
  rnaturalearth::ne_countries(returnclass = 'sf')
```

#Plot 

#Bounding box#
```{r}
africa_bbox <- c(-14.414063, -37.996163, 
                 53.613281, 27.994401)
xlim_af <- c(africa_bbox[1], africa_bbox[3])
ylim_af <- c(africa_bbox[2], africa_bbox[4])
```

#plot#
```{r}
ggplot(baseMap) +
  geom_sf() +
  geom_sf(fill = "#7fcdbb", colour = "#7fcdbb",
          data = nhm, show.legend = FALSE) +
  geom_sf(fill = "#2c7fb8", colour = "#2c7fb8",
          data = iucn2, show.legend = FALSE) +
  coord_sf(xlim = xlim_af, ylim = ylim_af, expand = TRUE)

```

#Create a convex hull instead for the IUCN map#
```{r}
newIUCN <- st_convex_hull(iucn2)
```

#Extract points that fall within the convex hull#
```{r}
overlap <- st_intersection(nhm, newIUCN)
```

#plot#
```{r}
ggplot(baseMap) +
  geom_sf() +
  geom_sf(fill = "#7fcdbb", colour = "#7fcdbb",
          data = nhm, show.legend = FALSE) +
  geom_sf(alpha = 0.5, fill = "#2c7fb8", colour = "#2c7fb8",
          data = newIUCN, show.legend = FALSE) +
  geom_sf(fill = "red", colour = "red",
          data = overlap, show.legend = FALSE) +
  coord_sf(xlim = xlim_af, ylim = ylim_af, expand = TRUE)
```

#Extract just the ID column from nhm data#
```{r}
nhm_simple <- nhm %>%
  dplyr::select(ID, Date, Extent_km, Cer) %>%
  st_drop_geometry()
```

#The overlap dataframe only contains the points that are overlapping with IUCN#
#So add a column of 1s for overlap score, then extract just the IDs and merge#
```{r}
overlap_simple <- overlap %>%
  mutate(overlap = rep(1, n())) %>%
  dplyr::select(ID, Date, Extent_km, Cer, overlap) %>%
  st_drop_geometry()
overlaps <- left_join(nhm_simple, overlap_simple)

```

#Add 0s for 0 overlap to replace NAs#
```{r}
overlaps$overlap[is.na(overlaps$overlap)] <- 0
```

#Look at the data#
```{r}
overlaps
```

#write file#
```{r}
write.csv(overlaps,"overlaps")
```