---
title: "Logistic general linear model"
author: "Alice Armstrong"
date: "2023-03-10"
output: html_document
---

This script performs a general linear model with binomial errors to measure the effects of human population size, density and land cover type on overlap between historic (museum specimen data) and present-day (IUCN data) C.niloticus distributions.

#Load packages#
```{r}
library(tidyverse)
library(car)
library(ggplot2)
library(patchwork)
library(rsq)
```
#load data#
```{r}
landuse<-read.csv("overlaps_landuse.csv")
hpd<-read.csv("overlaps_hpd.csv")
overlaps<-read.csv("overlaps")
```

#merge columns#
```{r}
hpd_land<-merge(landuse,hpd)
merged<-merge(overlaps,hpd_land)
```

#primary forest GLMs#
```{r}
primf1 <- glm(overlap ~ primf_1850,family = quasibinomial(link = "logit"),data = merged)
primf2 <- glm(overlap ~ primf_1900,family = quasibinomial(link = "logit"),data = merged)
primf3 <- glm(overlap ~ primf_1950,family = quasibinomial(link = "logit"),data = merged)
primf4 <- glm(overlap ~ primf_2000,family = quasibinomial(link = "logit"),data = merged)
```

#Summarise results and look at pseudo-r squared values#
```{r}
summary(primf1)
anova_res <- anova(primf1, test = 'F')
r<-rsq(primf1)

summary(primf2)
anova_res <- anova(primf2, test = 'F')
r<-rsq(primf2)

summary(primf3)
anova_res <- anova(primf3, test = 'F')
r<-rsq(primf3)

summary(primf4)
anova_res <- anova(primf4, test = 'F')
r<-rsq(primf4)

```

#plot primary forest plots for results section#

#primary forest 1850#
```{r}
plot_primf_1850<-ggplot(merged, aes(x=primf_1850, y=overlap)) + geom_point() +
  labs(y= "overlap", x = "primary forest ")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#primary forest 1900#
```{r}
plot_primf_1900<-ggplot(merged, aes(x=primf_1900, y=overlap)) + geom_point() +
  labs(y= "overlap", x = "primary forest")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
             method.args = list(family=binomial))
```

#primary forest 1950#
```{r}
plot_primf_1950<-ggplot(merged, aes(x=primf_1900, y=overlap)) + geom_point() +
  labs(y= "overlap", x = " primary forest")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#primary-non forest GLMs#
```{r}
primn1<-glm(overlap ~ primn_1850,family = quasibinomial(link = "logit"),data = merged)
primn2<-glm(overlap ~ primn_1900,family = quasibinomial(link = "logit"),data = merged)
primn3<-glm(overlap ~ primn_1950,family = quasibinomial(link = "logit"),data = merged)
primn4<-glm(overlap ~ primn_2000,family = quasibinomial(link = "logit"),data = merged)

```

#Summarise results and look at F stat and pseudo-r squared values#
```{r}
summary(primn1)
anova_res <- anova(primn1, test = 'F')
r<-rsq(primn1)

summary(primn2)
anova_res <- anova(primn2, test = 'F')
r<-rsq(primn2)

summary(primn3)
anova_res <- anova(primn3, test = 'F')
r<-rsq(primn3)

summary(primn4)
anova_res <- anova(primn4, test = 'F')
r<-rsq(primn4)
```

#plot primary non forest plots for results section#

#primary non forest 1850#
```{r}
plot_primn_1850<-ggplot(merged, aes(x=primn_1850, y=overlap)) + geom_point() +
  labs(y= "overlap", x = "primary non-forest")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#primary non forest 1900#
```{r}
plot_primn_1900<-ggplot(merged, aes(x=primn_1900, y=overlap)) + geom_point() +
  labs(y="overlap", x = "primary non-forest")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#primary non forest 1950#
```{r}
plot_primn_1950<-ggplot(merged, aes(x=primn_1950, y=overlap)) + geom_point() +
  labs(y= "overlap", x = " primary non-forest")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#primary non forest 2000#
```{r}
plot_primn_2000<-ggplot(merged, aes(x=primn_2000, y=overlap)) + geom_point() +
  labs(y= "% overlap", x = "primary non-forest")+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#urban land cover GLMs#
```{r}
urban1<-glm(overlap ~ urban_1850, family = quasibinomial(link = "logit"),data = merged)
urban2<-glm(overlap ~ urban_1900, family = quasibinomial(link = "logit"),data = merged)
urban3<-glm(overlap ~ urban_1950, family = quasibinomial(link = "logit"),data = merged)
urban4<-glm(overlap ~ urban_2000, family = quasibinomial(link = "logit"),data = merged)
```
#Summarise results and look at F stat and pseudo-r squared values#
```{r}
summary(urban1)
anova_res <- anova(urban1, test = 'F')
r<-rsq(urban1)

summary(urban2)
anova_res <- anova(urban2, test = 'F')
r<-rsq(urban2)

summary(urban3)
anova_res <- anova(urban3, test = 'F')
r<-rsq(urban3)

summary(urban4)
anova_res <- anova(urban4, test = 'F')
r<-rsq(urban4)
```

#plot urban for results section#

#urban 1850#
```{r}
plot_urban_1850<-ggplot(merged, aes(x=urban_1850, y=overlap)) + geom_point() +
  labs(y= "overlap", x = "urban")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#urban 1900# 
```{r}
plot_urban_1900<-ggplot(merged, aes(x=urban_1900, y=overlap)) + geom_point() +
  labs(y= "overlap", x = "urban")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```

#urban 1950#
```{r}
plot_urban_1950<-ggplot(merged, aes(x=urban_1950, y=overlap)) + geom_point() +
  labs(y= "overlap", x = "urban")+
  scale_y_continuous(breaks = c(0, 1), labels = c("No", "Yes"))+
  stat_smooth(method="glm", color="blue", se=FALSE,
              method.args = list(family=binomial))
```


#make a multiplot of land cover types#
```{r}
primn_plots<-plot_primn_1850+plot_primn_1900+plot_primn_1950+
  plot_primf_1850+plot_primf_1900+plot_primf_1950+
  plot_urban_1850+plot_urban_1900+plot_urban_1950

primn_plots + plot_annotation (tag_levels= 'i')
```


#-------------------population size and density GLMs----------------------------#


#population count#
```{r}
popc1<-glm(overlap~popc_1850,family = quasibinomial(link = "logit"),data = merged)
popc2<-glm(overlap~popc_1900,family = quasibinomial(link = "logit"),data = merged)
popc3<-glm(overlap~popc_1950,family = quasibinomial(link = "logit"),data = merged)
popc4<-glm(overlap~popc_2000,family = quasibinomial(link = "logit"),data = merged)
```

#Summarise results and look at F stat and pseudo-r squared values#
```{r}
summary(popc1)
anova_res <- anova(popc1, test = 'F')
r<-rsq(popc1)

summary(popc2)
anova_res <- anova(popc2, test = 'F')
r<-rsq(popc2)

summary(popc3)
anova_res <- anova(popc3, test = 'F')
r<-rsq(popc3)

summary(popc4)
anova_res <- anova(popc4, test = 'F')
r<-rsq(popc4)
```


#population density#
```{r}
popd1<-glm(overlap~popd_1850, family = quasibinomial(link = "logit"),data = merged)
popd2<-glm(overlap~popd_1900, family = quasibinomial(link = "logit"),data = merged)
popd3<-glm(overlap~popd_1950, family = quasibinomial(link = "logit"),data = merged)
popd4<-glm(overlap~popd_2000, family = quasibinomial(link = "logit"),data = merged)
```
#Summarise results and look at F stat and pseudo-r squared values#
```{r}
summary(popd1)
anova_res <- anova(popd1, test = 'F')
r<-rsq(popd1)

summary(popd2)
anova_res <- anova(popd2, test = 'F')
r<-rsq(popd2)

summary(popd3)
anova_res <- anova(popd3, test = 'F')
r<-rsq(popd3)

summary(popd4)
anova_res <- anova(popd4, test = 'F')
r<-rsq(popd4)

```

#rural population count#
```{r}
rurc1<-glm(overlap~rurc_1850, family= quasibinomial(link = "logit"),data = merged)
rurc2<-glm(overlap~rurc_1900, family= quasibinomial(link = "logit"),data = merged)
rurc3<-glm(overlap~rurc_1950, family= quasibinomial(link = "logit"),data = merged)
rurc4<-glm(overlap~rurc_2000, family= quasibinomial(link = "logit"),data = merged)
```
#Summarise results and look at F stat and pseudo-r squared values#
```{r}
summary(rurc1)
anova_res <- anova(rurc1, test = 'F')
r<-rsq(rurc1)

summary(rurc2)
anova_res <- anova(rurc2, test = 'F')
r<-rsq(rurc2)

summary(rurc3)
anova_res <- anova(rurc3, test = 'F')
r<-rsq(rurc3)

summary(rurc4)
anova_res <- anova(rurc4, test = 'F')
r<-rsq(rurc4)
```

#urban population count#
```{r}
urbc1<-glm(overlap~urbc_1850,family = quasibinomial(link = "logit"),data = merged)
urbc2<-glm(overlap~urbc_1900,family = quasibinomial(link = "logit"),data = merged)
urbc3<-glm(overlap~urbc_1950,family = quasibinomial(link = "logit"),data = merged)
urbc4<-glm(overlap~urbc_2000,family = quasibinomial(link = "logit"),data = merged)
```

#Summarise results and look at F stat and pseudo-r squared values#
```{r}
summary(urbc1)
anova_res <- anova(urbc1, test = 'F')
r<-rsq(urbc1)

summary(urbc2)
anova_res <- anova(urbc2, test = 'F')
r<-rsq(urbc2)

summary(urbc3)
anova_res <- anova(urbc3, test = 'F')
r<-rsq(urbc3)

summary(urbc4)
anova_res <- anova(urbc4, test = 'F')
r<-rsq(urbc4)

```